---
title: "Description of the data sets"
format: html
editor: visual
---

# Preparation

```{r}
#| message: FALSE
here::i_am("project_datamanagement.Rproj")
library(here)
library(vroom) 
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Links

## Link to the githib project

<https://github.com/inesdamiche/project_datamanagement.git>

## Links to the sources

World Bank Database We got the data by generating it through the World Bank Database by selecting all the countries , 2021 for the year and our variables of interest (Pourcentage of Women in Parliament , Female Labour Participation and Labour Participation, Annual GDP growth ) and by clicking on "Download options": [https://databank.worldbank.org/source/world-](https://databank.worldbank.org/source/world-development-indicators){.uri}[development](https://databank.worldbank.org/source/world-development-indicators){.uri}[-indicators](https://databank.worldbank.org/source/world-development-indicators){.uri}

We did the same thing for the meta-data (to categorise the countries by income group and regions): <https://databank.worldbank.org/source/world-development-indicators>

For the Human Development Composite Indices : <https://hdr.undp.org/data-center/documentation-and-downloads> We selected : "Human development Index and components" " Inequality-adjusted Human Development Index" "Gender development Index" "Planetary Pressure-adjusted HDI"

# Short description of the sources (500 words)

# Tables with basic information

## Looading the datasets

```{r}
#| message: FALSE
metadata <- vroom(here("data csv/countries_metadata.csv")) 
gender_dev <- vroom(here("data csv/dvlpmt_gender_index.csv"))
planet_HDI <- vroom(here("data csv/HDI_planet.csv"))
HDI <- vroom(here("data csv/HDI.csv"))
inequalities <- vroom(here("data csv/inequalities.csv"))
world_dev <- vroom(here("data csv/World Development Indicators - WB.csv"))
```

## Tables

```{r}
metadata |> 
  summarise("nb of rows in metadata" = nrow(metadata), "nb of columns in metadata"   = ncol(metadata))|>
  knitr::kable() 

gender_dev |> 
  summarise("nb of rows in gender_dev" = nrow(gender_dev), "nb of columns in metadata"   = ncol(gender_dev))|>
  knitr::kable()

HDI |> 
  summarise("nb of rows in HDI" = nrow(HDI), "nb of columns in HDI"   = ncol(HDI))|>
  knitr::kable()

inequalities |> 
  summarise("nb of rows in inequalities" = nrow(inequalities), "nb of columns in inequalities" = ncol(inequalities))|>
  knitr::kable()

planet_HDI |> 
  summarise("nb of rows in planet_HDI" = nrow(planet_HDI), "nb of columns in planet_HDI" = ncol(planet_HDI))|>
  knitr::kable()

#attention ici il faut remplacer .. par des NA 
world_dev |> 
  summarise("nb of rows in world_dev " = nrow(world_dev), "nb of columns in world_dev" = ncol(world_dev))|>
  knitr::kable()

```

# Short description of each variable

# Description of the data joining, cleaning, etc

expliquer comment on a réglé le problème des noms de pays différents

# Description of the research question (500 words)

# Graphical representation of the main variable (women in parliament)
