---
title: "Description of the data sets"
format: html
editor: visual
---

# Preparation 
```{r}
#| message: FALSE
here::i_am("project_datamanagement.Rproj")
library(here)
library(vroom) 
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Links

## Link to the githib project

<https://github.com/inesdamiche/project_datamanagement.git>

## Links to the sources 


# Short description of the sources (500 words)


# Tables with basic information 

## Looading the datasets  
```{r}
metadata <- vroom(here("data csv/countries_metadata.csv")) 
gender_dev <- vroom(here("data csv/dvlpmt_gender_index.csv"))
planet_HDI <- vroom(here("data csv/HDI_planet.csv"))
HDI <- vroom(here("data csv/HDI.csv"))
inequalities <- vroom(here("data csv/inequalities.csv"))
world_dev <- vroom(here("data csv/World Development Indicators - WB.csv"))
```
## Tables
```{r}

```


# Short description of each variable 

# Description of the data joining, cleaning, etc
expliquer comment on a réglé le problème des noms de pays différents 

# Description of the research question (500 words)

# Graphical representation of the main variable (women in parliament)





