---
title: "Description of the data sets"
format: html
---

# Preparation

```{r}
#| message: FALSE
here::i_am("project_datamanagement.Rproj")
library(here)
library(vroom) 
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Links

## Link to the github project

<https://github.com/inesdamiche/project_datamanagement.git>

## Links to the sources

### For the World Bank Database

We got the data by generating it through the World Bank Database by selecting all the countries , 2021 for the year and our variables of interest (Percentage of Women in Parliament , Female Labour Participation and Labour Participation, Annual GDP growth ) and by clicking on "Download options": <https://databank.worldbank.org/source/world-development-indicators>

We did the same thing for the meta-data (to categorise the countries by income group and regions): <https://databank.worldbank.org/source/world-development-indicators>

### For the Human Development Composite Indices

<https://hdr.undp.org/data-center/documentation-and-downloads> We selected : "Human development Index and components" " Inequality-adjusted Human Development Index" "Gender development Index" "Planetary Pressure-adjusted HDI"

The meta-data for these databases is taken from the UN's Statistics Division website : we downloaded the CSV version <https://unstats.un.org/unsd/methodology/m49/overview/>

# Short description of the sources

## World Development Indicators

The World Development Indicators are compiled by the **World Bank**. There are "a compilation of relevant, high quality, and internationally comparable statistics about global development and the fight against poverty".

### Indicator: Proportion of seats held by women in national parliaments (%)

The data is compiled by the **Inter-Parliamentary Union (IPU)** (<https://ipu.org/>) based on information provided by National Parliaments. The data is compiled every year.

### Indicator: Women Business and the Law Index Score (scale 1-100)

The data is compiled every year by the **World Bank** (<https://wbl.worldbank.org/>).

### Indicator: Military expenditure (% of GDP)

The data is compiled every year by the **Stockholm International Peace Research Institute (SIPRI)** (<https://www.sipri.org/databases>).

Because of the differences in definitions and the difficulty in verifying the accuracy and completeness of data, data on military expenditures are not always comparable across countries. However, SIPRI puts a high priority on ensuring that the data series for each country is comparable over time.

### Indicator: Labor force participation rate, total (% of total population ages 15+) (modeled ILO estimate)

The data is compiled by the **International Labor Organization** (<https://ilostat.ilo.org/data/>) from labor force surveys, censuses, and establishment censuses and surveys. It is part of the "ILO modeled estimates database". The data is compiled every year.

### Indicator: Labor force participation rate, female (% of female population ages 15-64) (modeled ILO estimate)

The data is compiled in the same way as for the previous indicator.

### Indicator: GDP growth (annual %)

The data come from **the World Bank** national accounts data and **OECD** National Accounts files. It is compiled every year.

## Human Development Indices

The Human Development Indices are compiled by the **UNDP** (<https://www.undp.org>). The Human Development Indices are "a measure of average achievement in key dimensions of human development".

### Indicator: planetary pressures-adjusted Human Development Index (HDI, PHDI, adjustment factor for planetary pressures, Carbon dioxide emissions per capita (production), Carbon dioxide emissions (production) index, Material footprint per capita, Material footprint index)

From this dataset compiled by the **UNDP**, we chose to only use the indicator "Planetary pressures-adjusted HDI (PHDI)" because we think it is the more relevant and concise one. The PHDI is compiled by the UNDP. The PHDI is computed as the product of the HDI and the "adjustment factor for planetary pressures". This factor is computed thanks to the **United Nations Environment Programme (2022)** and the **Global Carbon Project (2022)**.

### Indicator: Human development index and its components (HDI, Life expectancy at birth, expected years of schooling, mean years of schooling, gross national income per capita, GNI per capita rank minus HDI rank, HDI rank)

From this dataset compiled by the **UNDP**, we chose to use the indicator "HDI" because we think it is the more relevant and concise one. To compute the HDI, the PNUD uses different indicators:

-   Life expectancy at birth: **UNDESA (2022)**

-   Expected years of schooling: **UNESCO Institute for Statistics (2022), ICF Macro Demographic and Health Surveys, United Nations Children's Fund (UNICEF) Multiple Indicator Cluster Surveys and OECD (2022)**

-   Mean years of schooling: **UNESCO Institute for Statistics (2022), Barro and Lee (2018), ICF Macro Demographic and Health Surveys, UNICEF Multiple Indicator Cluster Surveys and OECD (2022)**

-   GNI per capita: **World Bank (2022), IMF (2022) and United Nations Statistics Division (2022)**

### Indicator: Gender development index

From this dataset, we chose to only use the indicator "GDI" because we think it is the more relevant and concise one. GDI is computed thanks to different indicators:

-   Life expectancy at birth: **UNDESA (2022)**

-   Expected years of schooling: **UNESCO Institute for Statistics (2022), ICF Macro Demographic and Health Surveys, United Nations Children's Fund (UNICEF) Multiple Indicator Cluster Surveys and OECD (2022)**

-   Mean years of schooling for adults ages 25 andcolder: **UNESCO Institute for Statistics (2022), Barro and Lee (2018), ICF Macro Demographic and Health Surveys, UNICEF's Multiple Indicator, Cluster Surveys and OECD (2022)**

-   Estimated earned income: **Human Development Report Office estimates based on female and male shares of the economically active population, the ratio of the female to male wage in all sectors and gross national income in 2017 purchasing power parity (PPP) terms, and female and male shares of population from ILO (2022), UNDESA (2019), World Bank (2022), United Nations Statistics Division (2022) and IMF (2022)**

### Indicator: Inequality-adjusted Human Development Index (Coefficient of human inequality, Inequality in life expectancy, Inequality-adjusted life expectancy index, Inequality in education, Inequality-adjusted education index, Inequality in income, Inequality-adjusted income index, Income shares held by (%), Gini coefficient)

From this dataset, we chose to only use the indicator "Gini coefficient" because we think it is the more relevant and concise one. The Gini coefficient comes from the **World Inequality Database (2022)** (<https://wid.world>).

# Tables with basic information

## Loading the datasets

```{r}
#| message : false
metadata <- vroom(here("data csv/countries_metadata.csv")) 
gender_dev <- vroom(here("data csv/dvlpmt_gender_index.csv"))
planet_HDI <- vroom(here("data csv/HDI_planet.csv"))
HDI <- vroom(here("data csv/HDI.csv"))
inequalities <- vroom(here("data csv/inequalities.csv"))
world_dev <- vroom(here("data csv/World Development Indicators - WB.csv"))
metadata_UNDP <- vroom(here("data csv/UNSD — Methodology.csv")) 
```

## Tables (before cleaning and joining)

```{r}
metadata |> 
  summarise("nb of rows in metadata" = nrow(metadata), "nb of columns in metadata"   = ncol(metadata))|>
  knitr::kable() 

metadata_UNDP |> 
  summarise("nb of rows in metadata UNDP" = nrow(metadata_UNDP), "nb of columns in metadata UNDP"   = ncol(metadata_UNDP))|>
  knitr::kable() 

gender_dev |> 
  summarise("nb of rows in gender_dev" = nrow(gender_dev), "nb of columns in metadata"   = ncol(gender_dev))|>
  knitr::kable()

HDI |> 
  summarise("nb of rows in HDI" = nrow(HDI), "nb of columns in HDI"   = ncol(HDI))|>
  knitr::kable()

inequalities |> 
  summarise("nb of rows in inequalities" = nrow(inequalities), "nb of columns in inequalities" = ncol(inequalities))|>
  knitr::kable()

planet_HDI |> 
  summarise("nb of rows in planet_HDI" = nrow(planet_HDI), "nb of columns in planet_HDI" = ncol(planet_HDI))|>
  knitr::kable()

world_dev |> 
  summarise("nb of rows in world_dev " = nrow(world_dev), "nb of columns in world_dev" = ncol(world_dev))|>
  knitr::kable()

```

# Short description of each variable

### Human Development Index

The Human Development Index (HDI) is a composite index that measures a country's average achievement in human development. It takes into account three dimensions: health, education, and standard of living. The HDI is calculated by taking the geometric mean of three normalized indices for each dimension: life expectancy at birth (health), mean years of schooling for adults aged 25 years, and expected years of schooling for children of school-entering age (education), and the gross national income per capita measured in logarithm (standard of living). (UNDP , 2023)

### Gender Development Index

The Gender Development Index (GDI) is a variant of the HDI which measures gender inequalities in the achievement of human development. It takes into account the same dimensions as HDI while considering the disparities between men and women. The indicators for each dimensions are : female and male life expectancy at birth (health) , female and male expected years of schooling for children and female and male mean years of schooling for adults ages 25 years and older (education) and female and male estimated earned income (standard of living/command over economic resources ).(UNDP , 2023)

### Gini Coefficient

The Gino coefficient aims to measures income inquality within a population. It takes a value from 0 (perfect equality) to 1 (perfect inequality). It illustrates to extent to which the income distribution deviates from perfect equality.

### Planetary Pressures- Adjusted Human Development Index

The Planetary Pressures-Adjusted Human Development Index ( PHDI) is also a variant of the HDI which discounts from the HDI the pressures on the planet. It is the level of HDI adjusted by carbon dioxide emissions per person (production-based) and material footprint per capita to account . With no pressures on the planet , it is equal to the HDI. It integrates environmental and sustainability concerns into the measurement of Human Development. (UNDP , 2023)

### Proportion of seats held by women in national parliaments(in %)

It is the percentage of parliamentary seats (single or lower chamber) held by women in a country.

### Labor force participation rate , female (% of female population ages 15+)(modeled ILO estimate) and Labor force participation rate , total (% of total population ages 15+)(modeled ILO estimate)

The Labor force participation rate is the proportion of the population ages 15 and older that are economically active.Here "economically active" concerns "all people who supply labor for the production of goods and services during a specified period". (World Bank , 2023).

The Female Labor force participation rate is the proportion of the female population ages 15 and older that are economically active

### GDP Growth (anuual %)

The Annual GDP Growth is the Aanual percentage growth rate of GDP at market prices based on constant local currency. According to the World Bank's (2023) definition , the GDP is the sum of gross value added by all resident producers in the economy plus any product taxes and minus any subsidies not included in the value of the products. It aims to measure economic growth.

### Women Business and the Law Index Score (scale 1-100)

The Women Business and the Law Index Score is an Index developed by the World Bank that measures on a scale from 1 to 100 how laws and regulations affect women's economic opportunity.It is computed by taking the average score of 9 indices (Mobility, Workplace, Pay, Marriage, Parenthood, Entrepreneurship, Assets and Pension ). (World Bank , 2023)

### Military expenditure (% of GDP)

The Military expenditure (% of GDP) includes the proportion in the GDP of all current and capital expenditures on the armed forces, including peacekeeping forces; defense ministries and other government agencies engaged in defense projects; paramilitary forces. It aims to measure a nation's prioritisation of defense spending relative to its overall economic activity. (World Bank , 2023)

# Description of the data joining and cleaning

## Data Cleaning

### World Development Indicators Dataset

The first step involved cleaning the World Development Indicators dataset in order to convert the missing values denoted as ",,", into NA values.In addition , we converted the data type of the value of each indicator from string to numeric, while making sure to change to decimal indicator from "," to "." for RStudio to correctly identify the value as decimal.

```{r}
# | message : false 
# Data Cleaning - World Development Indicators.
# We turn the rows that do not contains a values into NA value
Dev <- world_dev |>
  mutate(`2021 [YR2021]` = na_if(`2021 [YR2021]`, ",,"))
# We turn the values into numeric ones and make sure that Rstudio takes into account that the decimal separator is "," and not "."
Dev <- Dev|>
  mutate(`2021 [YR2021]` = as.numeric(gsub(",",".", `2021 [YR2021]`)))
```

### UNDP Databases

For the UNDP Databases ( Human Development Indices) , this step involved selecting the columns of the variables we are interested in , changing the decimal indicators from "," to "." and renaming the columns in order to have uniform names across all databases.

```{r}
# UNDP Database
#we select columns we are interested in and turn the values into numeric ones 
HDI<-
  HDI|>
  select(`Table 1. Human Development Index and its components`,...3)|>
  rename(`Country Name`=`Table 1. Human Development Index and its components`,HDI_val=...3)
HDI <- HDI|>
  mutate(HDI_val = as.numeric(gsub(",", ".", HDI_val)))

#Gender Development Index
GDI<-gender_dev|>
  select(`Table 4. Gender Development Index`,...3)|>
  rename(`Country Name`=`Table 4. Gender Development Index`,GDI_val=...3)|>
   mutate(GDI_val = as.numeric(gsub(",", ".", GDI_val)))

# Gini Coefficient
Gini<-inequalities|>
  select(`Table 3. Inequality-adjusted Human Development Index`,...31)|>
  rename(`Country Name`=`Table 3. Inequality-adjusted Human Development Index`,Gini_coef=...31)|>
  mutate(Gini_coef = as.numeric(gsub(",", ".", Gini_coef)))

#PHDI
PHDI<-planet_HDI|>
  select(`Table 7. Planetary pressures-adjusted Human Development Index`,...5)|>
  rename(`Country Name`=`Table 7. Planetary pressures-adjusted Human Development Index`,PHDI=...5)|>
  mutate(PHDI = as.numeric(gsub(",", ".", PHDI)))

```

### Meta-data Databases

For the Meta-data Databases for both the World Bank's and the UNDP's databases , this step involved selecting the columns with the variables we are interested in and renaming them to ensure the uniformity of the variables' names across the datasets.

```{r}
# Metadata cleaning 
#World Bank
meta_WB<-metadata|>
  select(Code,`Table Name`,Region,`Income Group`)|>
  rename(`Country Name`=`Table Name`)
#UNDP
meta_UNDP<-metadata_UNDP|>
  select(`Country or Area`,`ISO-alpha3 Code`)|>
  rename(`Country Name`=`Country or Area`)
```

## Data Joining

### Meta-Data and Databases

For the UNDP datasets , we computed a function to simplify the joining process with the meta-data. The aim of the function is to perform a left join with the UNDP database and the meta-data database,renaming the columns names and deleting the country name from the final dataset in order to join them with the main dasaset ( World Development Indicators) after.

Because some countries' names were computed in the meta-data database, we computed them manually by creating a separate dataset with the country names (as in the UNDP datasets) and their country code and binded them to the initial UNDP meta-data database.

```{r}
#MetaData Joining - UNDP databases
#join UNDP data with their Country Code

# Function in order to perform a "left-join" for each databases
country_code<-function(dataset1,dataset2){
  dataset1<-left_join(dataset1, dataset2, join_by(`Country Name`))|>
    rename(`Country Code`=`ISO-alpha3 Code`)|>
    drop_na(`Country Code`)|>
    select(-`Country Name`)
  
}

# we create a new metadata dataframe to join the country code for the 12 countries that couldn't be joined

other <- tibble(`Country Name`  = c("United States", "United Kingdom", "Tanzania (United Republic of)", "Palestine, State of", "Netherlands","Moldova (Republic of)",	"Korea (Republic of)" , "Korea (Democratic People's Rep. of)" , "Hong Kong, China (SAR)",
"Eswatini (Kingdom of)" , "Côte d'Ivoire" , "Congo (Democratic Republic of the)"),
  `ISO-alpha3 Code` = c("USA","GBR","TZA","PSE","NLD", "MDA","KOR","PRK","HKG","SWZ","CIV","COD"))
meta_UNDP<-rbind(meta_UNDP,other)


HDI<-country_code(HDI,meta_UNDP)
GDI<-country_code(GDI,meta_UNDP)
Gini<-country_code(Gini,meta_UNDP)
PHDI<-country_code(PHDI,meta_UNDP)
```

For the World Bank datase , we joined the meta-data with region and the income groups by country code to the initial database.

```{r}
#MetaData Joining - WB databases
meta_WB<-meta_WB|>
  select(-`Country Name`)|>
  rename(`Country Code`=`Code`)
Dev<-left_join(Dev, meta_WB, join_by(`Country Code`))

```

As a last step , we joined the UNDP databases to the World Development Indicators' one.

```{r}
# Databases Joining
Dev <-inner_join(Dev,GDI,join_by(`Country Code`))
Dev <-inner_join(Dev,HDI,join_by(`Country Code`))
Dev <-inner_join(Dev,Gini,join_by(`Country Code`))

# Final Dataset
dataset <- inner_join(Dev, PHDI, join_by(`Country Code`))


```

### Final Dataset

Finally , we created the final dataset by using the pivot_wider() function in order to transform the world development indicators into columns to have separate variables. Then, we eliminated all the countries that had at least one NA .

```{r}
# Turn the Indicators from the World Bank Databases into Columns 
dataset<-dataset|>
  select(-`Series Code`)|>
  pivot_wider(names_from = `Series Name`,values_from = `2021 [YR2021]`)

# Elimination of all the NA Values 
dataset<-na.omit(dataset)
```

# Description of the research question

The 2023 Agenda for Sustainable Development puts a strong emphasis on advancing gender equality and the empowerment of women in leadership and decision‑making roles within the framework of SDG 5: Gender Equality. Indeed, with women constituting almost half of every country's population, their inclusion in the decision-making process is key to ensuring more representation in politics.

Beyond transforming social norms and ensuring diversity, research has also shown that the involvement of women in politics could lead to shifts in policy priorities by positively influencing legislation regarding social justice, education, and peace. For instance, O'Brien and Piscopo (2018) demonstrated that female representatives can diversify legislative agendas in favor of vulnerable groups, while Ramstetter and Habersack (2019) found that within the European Parliament, women were notably more supportive of environmental legislation than men. This suggests that increasing the number of women in the legislative process could have a positive impact on a country's development by shifting public policy orientations.

Our study aims to explore the correlation between the proportion of women occupying parliamentary seats and the implementation of public policies. We seek to analyze the interplay between women in parliament and various development and socio-economic indicators to identify recurring patterns. Thus, **we will investigate the extent to which the proportion of seats held by women in national parliaments influences public policy outcomes** by conducting an analysis across countries at the world level, the regional level, and within income groups to better understand the disparities and nuances within similar countries in terms of development.

To account for different dimensions of a country's development and policy priorities, we will consider several indicators offering different lenses through which we can assess the influence of women in parliaments on public policies.

We will first analyze basic development indicators such as the **Human Development Index (HDI) and the Annual GDP growth** to assess whether women's participation in decision-making correlates positively with human and economic development. Given the existing correlation between higher development and gender equality, we will also conduct analyses within regions and income groups to see if there are any patterns.

In addition, we will also introduce other economic variables such as the **Total Labor Participation Rate and the Female Participation Rate** to assess the correlation with employment and economic inclusion.

Furthermore, this study also aims to investigate the outcome of public policies regarding gender development with indicators such as the **Gender Development Index (GDI) and the Women Business and the Law Index Score**, which reflects legal frameworks supporting women in business and law.

To assess potential shifts in policy orientation towards social justice and equality rather than economic growth and military spending, we will incorporate the **Military Expenditure (% of GDP)** indicator and the **Gini Coefficient ,** which provides a measure for policies regarding income inequality and wealth distribution.

\
Finally, we will evaluate the relationship between women parliamentarians and the priority given to environmental issues through the incorporation of the **Planetary Pressures-Adjusted Human Development Index (PHDI).**

# Graphical representation of the main variable (women in parliament)

## Distribution of the proportion of seats held by women in national parliaments (in percentage) ...

### ...In the World

```{r}
dataset |>
  ggplot(aes(x = `Proportion of seats held by women in national parliaments (%)`)) +
  geom_density(fill = "orange") +
  labs(title = "Distribution of Proportion of Seats Held by Women in National Parliaments",
       x = "Proportion of Seats Held by Women (%)",
       y = "Density") 

```

### ... By Income Group

```{r}
dataset |>
  ggplot(aes(x = `Proportion of seats held by women in national parliaments (%)`)) +
  geom_density(fill = "orange") +
  labs(title = "Distribution of Proportion of Seats Held by Women in National Parliaments",
       x = "Proportion of Seats Held by Women (%)",
       y = "Density") +
  facet_wrap(~`Income Group`)
```

### ... By Region

```{r}
dataset |>
  ggplot(aes(x = `Proportion of seats held by women in national parliaments (%)`)) +
  geom_density(fill = "orange") +
  labs(title = "Distribution of Proportion of Seats Held by Women in National Parliaments",
       x = "Proportion of Seats Held by Women (%)",
       y = "Density") +
  facet_wrap(~`Region`)
```

### Sources

United Nations Development Programme (2023). Human Development Index. \[online\] United Nations Development Programme. Available at: https://hdr.undp.org/data-center/human-development-index#/indicies/HDI.

United Nations Development Programme (2023). Gender Development Index. \[online\] hdr.undp.org. Available at: https://hdr.undp.org/gender-development-index#/indicies/GDI.

United Nations Development Programme (2023). Planetary pressures--adjusted Human Development Index. \[online\] hdr.undp.org. Available at: https://hdr.undp.org/planetary-pressures-adjusted-human-development-index#/indicies/PHDI.

World Bank (2023). GDP Growth (annual %) \| Data. \[online\] Worldbank.org. Available at: https://data.worldbank.org/indicator/ny.gdp.mktp.kd.zg

World Bank (2023).. Women Business and the Law Index Score (scale 1-100) \| Data. \[online\] Available at: https://data.worldbank.org/indicator/SG.LAW.INDX. The World Bank (2023). Military expenditure (% of GDP) \| Data. \[online\] Worldbank.org. Available at:https://data.worldbank.org/indicator/MS.MIL.XPND.GD.ZS.

‌O\'Brien, D.Z. and Piscopo, J.M. (2018). The Impact of Women in Parliament. *The Palgrave Handbook of Women\'s Political Rights*, pp.53--72.

Ramstetter, L. and Habersack, F. (2019). Do Women Make a difference? Analysing Environmental Attitudes and Actions of Members of the European Parliament. *Environmental Politics*, 29(6), pp.1--22.
